Viscous Erosion in Porous Media

### 2018-03-23 docs/methods/erosion_methods.tex
NM

Figure~\ref{shrink_intface}(a) shows that the spacing between successive interfaces becomes larger with time, suggesting that the overall shear stress grows as the body vanishes. This observation is consistent with the well-known 3D Stokes drag law, in which stress, $\tau \sim \mu \umax / a$, scales inversely proportional to body-scale $a$ (e.g.~for a sphere, $a$ is the radius). This law must be modified in two dimensions, though. The modification is made non-trivial by subtleties of the 2D Stokes paradox, but can be correctly deduced by noting that, as the body vanishes, its drag must be consistent with slender-body theory \cite{poz1997, bouzarth2007epicyclic, camassa2008trajectory, MooreJFM2012, zhao2015transportation}. 
For a slender body in flow perpendicular to its long-axis $\ell$, the drag is proportional to
$\mu \umax \ell/ \log(\ell/a)$, where $a \ll \ell $ is a length scale of the cross-section. This relationship implies that the stress-per-unit-span on a 2D cross-section can be estimated as
\begin{align}
  \label{eqn:stresslaw}
  \tau \sim \frac{\mu \umax}{a \log(\ell/a)} \qquad \text{for } a \ll \ell \, .
\end{align}
In interpreting our 2D simulations, we will treat the span, $\ell$, as
fixed. Recall that $L$ denotes the total perimeter of the 2D cross-section. Since $L$ and $a$ each represent a characteristic scale of the cross-section, they can often be interchanged in scaling analysis.

@@

	Another important quantity that can be used to test our simulations is the drag exerted by the body as it erodes. Figure~\ref{area_drag}(b) shows the horizontal drag, $F_D$, calculated in the simulations (black curve) using Eq.~\eqref{eqn:drag}. In the context of slender-body theory, $F_D$ should be interpreted as drag-per-unit-span. As seen in Fig.~\ref{area_drag}(b), the drag decreases gradually at first and then abruptly tends to zero in the very final moments. This behavior, though perhaps surprising, is consistent with the fact that drag vanishes {\em logarithmically} with body-size in 2D Stokes flow. Equation \eqref{eqn:drag} also allows us to calculate the separate contributions of pressure and viscous drag, and we show these with the blue and green curves in Fig.~\ref{area_drag}(b). Initially, pressure and viscous drag are comparable, with the circular geometry having a slightly higher level of pressure drag. As the body erodes and assumes a more slender form, though, the pressure drag decreases much more rapidly, leaving viscous drag to dominate. This behavior is associated with the tendency towards a drag-minimizing profile, which will be discussed in Section \ref{LimitingShape}.

>>>>

Figure~\ref{shrink_intface}(a) shows that the spacing between successive interfaces becomes larger with time, suggesting that the overall shear stress grows as the body vanishes. This observation is consistent with the well-known 3D Stokes drag law, in which stress, $\tau \sim \mu \umax / a$, scales inversely proportional to body-scale $a$ (e.g.~for a sphere, $a$ is the radius). This law must be modified in two dimensions, though. The modification is made non-trivial by subtleties of the 2D Stokes paradox, but can be correctly deduced by noting that, as the body vanishes, its drag must be consistent with slender-body theory \cite{poz1997, bouzarth2007epicyclic, camassa2008trajectory, MooreJFM2012, zhao2015transportation}. 
For a slender body in flow perpendicular to its long-axis $\ell$, the drag is proportional to
$\mu \umax \ell/ \log(\ell/a)$, where $a \ll \ell $ is a length scale of the cross-section. This law implies that the stress-per-unit-span on a 2D cross-section is approximately
\begin{align}
  \label{eqn:stresslaw}
  \tau \sim \frac{\mu \umax}{a \log(\ell/a)} \qquad \text{for } a \ll \ell \, .
\end{align}
In interpreting our 2D simulations, we will treat the span, $\ell$, as
fixed. Recall that $L$ denotes the total perimeter of the 2D cross-section. Since $L$ and $a$ each represent a characteristic scale of the cross-section, they can often be interchanged in scaling analysis.

@@

	As the body shrinks in accordance with \eqref{area_predict}, the drag that it exerts decreases at a related rate, which offers a second check on our numerics. In Fig.~\ref{area_drag}(b), we show the horizontal drag, $F_D$, on the simulated body as it erodes. The drag is calculated using ...

	Figure~\ref{area_drag}(b) shows the horizontal drag, $F_D$, measured in the simulations (black curve) using Eq.~\eqref{eqn:drag}. In the context of slender-body theory, $F_D$ should be interpreted as drag-per-unit-span. As seen in Fig.~\ref{area_drag}(b), the drag decreases gradually at first and then abruptly tends to zero in the very final moments. This behavior, though perhaps surprising, is consistent with the fact that drag vanishes {\em logarithmically} with body-size in 2D Stokes flow. Equation \eqref{eqn:drag} also allows us to calculate the separate contributions of pressure and viscous drag, and we show these with the blue and green curves in Fig.~\ref{area_drag}(b). Initially, pressure and viscous drag are comparable, with the circular geometry having a slightly higher level of pressure drag. As the body erodes and assumes a more slender form, though, the pressure drag decreases much more rapidly, leaving viscous drag to dominate. This behavior is associated with the tendency towards a drag-minimizing profile, which will be discussed in Section \ref{LimitingShape}.
